Pool Model

### 2025-08-28 pool_model.tex
write introductory part for spatial effects

As described above, the proposed model operates with the total number of the bacterial counts for different species and states in the system not accounting for their spatial distributions.
Assuming that the bacterial system is uniformly distributed (sometimes also called 'well-stirred') we can simplify the problem to a system of \acp{ode}.
However, in reality spatial effects can play a significant role for population survival forcing some species develop competitive mechanisms such as cell adhesion to surfaces in the areas rich with nutrients~\cite{htuson_bacteriasurface_2013}.
Another mechanism is the motility of bacteria, e.g., the flagellates bacteria as \textit{E. coli} or \textit{B. subtilis} can experience the chemotactic motion of driven by oxygen concentrations~\cite{decoene_microscopic_2011}.
The bacterial social interactions such as releasing the toxins or public goods can also significantly affect the spatial structure of the community~\cite{blanchard_bacterial_2015}.
These effects can only be treated with a spatial model that is able to describe the observed effects.\\

@@

To test the spatial limitations of the described above approach consider the system consisting from two bacterial species that can be described by two internal states: lag and growth states. 
Both of these species grow using limiting resource pool $R^*$ with different growth constants $\alpha$ and different lengths of lag phase determined by parameter $\lambda$.
It has to be noted that the resource pool is normalized by its initial value, i. e., $R^* = R/R_0$.
Moreover, the species $A$ is able to release a certain antimicrobial component $I$ that shows inhibitory effect on species 2 with any harmful effect on itself (amensalism).
We assume that $L,G$ are measured in terms of overall volume and $R,I$ are measured as total amount.

>>>>

As described above, the proposed model operates with the total number of the bacterial counts for
different species and states in the system, not accounting for their spatial distributions.
Assuming that the bacterial system is uniformly distributed (sometimes also called 'well-stirred')
we expect to be able to simplify the problem and map it to a system of \acp{ode}.
However, in reality spatial effects can play a significant role for population survival forcing some
species to develop competitive mechanisms such as cell adhesion to surfaces in the areas rich with
nutritients~\cite{htuson_bacteriasurface_2013}.
Another mechanism is the motility of bacteria, e.g., the flagellates bacteria as \textit{E. coli} or
\textit{B. subtilis} can experience the chemotactic motion of driven by oxygen
concentrations~\cite{decoene_microscopic_2011}.
The bacterial social interactions such as releasing the toxins or public goods can also
significantly affect the spatial structure of the community~\cite{blanchard_bacterial_2015}.
These effects can only be treated with a spatial model that is able to describe the observed
effects.\\

@@

To test the spatial limitations of the described above approach consider the system consisting from
two bacterial species that can be described by two internal states: lag and growth states.
Both of these species grow using the same resource pool $R$ but with different growth constants
$\alpha$ and different lengths of lag phase $\lambda$.
Moreover, the species $A$ is able to release a certain antimicrobial component $I$ that shows an
inhibitory effect on species 2 (amensalism).
We assume that $L,G$ are measured in terms of the overall volume and $R,I$ are measured as total
amount.
